---
title: "Regression"
output: pdf_document
date: "2025-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(car)
library(lmtest)
library(sandwich)
library(dplyr)
library(readxl)

# Load data
data <- read_excel("wages.xlsx")
```



```{r}
# 1. Plot Wage against Age and evaluate whether a linear or quadratic model would better capture the relationship. 
ggplot(data, aes(x = Age, y = Wage)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", formula = y ~ x, col = "red", se = FALSE) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), col = "blue", se = FALSE) + 
  labs(title = "Wage vs Age",
       x = "Age",
       y = "Wage") +
  theme_minimal()

```

## A quadratic model would better capture the relationship since the points are going in a downward U-shape and not in a linear shape


```{r}
#Estimate a multiple regression model of Wage using Age and Education as independent (X) variables; assume a standard linear relationship between Wage and Age.
linear_model <- lm(Wage ~ Age + Educ, data = data)

# Summary of the model
summary(linear_model)
```

#we get a R-squared of 0.6088 and the most significant variable being education more than age, and age being impacted by education years the most

```{r}
#Estimate another multiple regression model of Wage using Age and Education as independent (X) variables; this time fit Age using a quadratic relationship. Verify your choice from part a. by comparing the distribution of residuals and the goodness of fit between the models in parts b and c.

quad_model <- lm(Wage ~ Age + I(Age^2) + Educ, data = data)
summary(quad_model)
```

#we can see that now with age having a quadratic relationship, the R-squared value went up in accuracy.

```{r}
#Use the appropriate model to predict hourly wages for someone with 16 years of education and age equal 30, 50, or 70.

new_data <- data.frame(Age = c(30, 50, 70), Educ = rep(16, 3))

predicted_wages <- predict(quad_model, newdata = new_data)
predicted_wages
```

#someone with 50 years of age will obtain the highest wages 

```{r}
#residuals and fitted values with our models
plot(fitted(linear_model), residuals(linear_model), 
     main="Residuals vs Fitted",
     xlab="Fitted values", ylab="Residuals")
abline(h=0, col="red")

plot(fitted(quad_model), residuals(quad_model), 
     main="Residuals vs Fitted",
     xlab="Fitted values", ylab="Residuals")
abline(h=0, col="red")
```

#after plotting residuals and fitted values we can confirm the quadratic model has more points closer to the line than on the outside like the linear model

```{r}
#Plot Rent against each of the three predictor variables and evaluate whether the relationship is best captured by a line or a curve. Identify variables that may benefit from a log-transformation 

data <- read_excel("AnnArbor.xlsx")

ggplot(data, aes(x = Beds, y = Rent)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Rent vs. Bedrooms")

ggplot(data, aes(x = Baths, y = Rent)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Rent vs. Bathrooms")
```

#both bedrooms and bathrooms both are going up and appear somewhat linear but with different jumps but bathrooms have a more nonlinear trends

```{r}
ggplot(data, aes(x = Sqft, y = Rent)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Rent vs. Square Footage")
```

#for the square feet it looks like it may a log transformation since it does have a curved pattern
```{r}
data$Log_Sqft <- log(data$Sqft)


model <- lm(Rent ~ Beds + Baths + Log_Sqft, data = data)

# Display model summary
summary(model)
```

#model has an r-squared of 0.8355 degrees meaning its a good fit and our t statistic showing it is statistically significant with most square feet being the most significant and beds too

```{r}
new_data <- data.frame(Beds = 3, Baths = 2, Log_Sqft = log(1600))

# Predict rent
predicted_rent <- predict(model, newdata = new_data)

# Print predicted rent
print(predicted_rent)
```

#the predicted rent is 1540.384 for a 3 beds 2 baths and 1600 squared feet
